---
title: "Dynamic Measurement of the Capacity of Federal Agencies"
subtitle: "PS 919: Professor Tahk"
author: "Blain Morin"
date: "12/12/2020"
output: pdf_document
bibliography: bib.bib
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

```{r}


library(knitr)
library(tidyverse)
library(rstan)
library(parallel)
library(pscl)
library(kableExtra)
library(shinystan)
library(ggcorrplot)
library(gridExtra)
library(ggridges)
library(RColorBrewer)

```


## Introduction

The capacity of a bureaucratic agency, or its resources and professionalization, play a key role in an agency’s “ability to induce residents, firms, and organizations to act in ways they would not in the absence of its regulatory and administrative presence” [@berch; @kurtz]. Measuring this “institutional power” of federal agencies is of interest to political scientists and policy researchers because it is an important input to governance outcomes [@berch]. This paper uses a dynamic latent variable approach to measure the capacity of federal agencies in the United States.

## Data

The ensuing analysis uses personnel data from all federal agencies from 1973 to 2019. The data set contains uniquely identified information (such as education level and salary) for individual federal employees. These data not only allow us to calculate aggregate measures (such as the median salary of an agency), but it also allows us to see when an agency exudes and absorbs employees to and from other agencies. The data come from the U.S. Office of Personnel Management and were made publicly available by BuzzFeed News [@buzz].     

We use the individual data create aggregate measures for each agency for each year. The main agency level variables related to capacity are:

* *Number of Employees*
* *Median Salary*
* *Percentage of Employees with a Master's Degree or Higher*
* *Average Length of Service*
* *Percentage of Employees Requisitioned from Other Agencies*
* *Percentage of Employees who Left the Agency*
* *Gini Coefficient*

## Exploratory Analysis



```{r}

fed = read_csv("fed_agency_capacity_autonomy.csv")

set.seed(9876)

df = fed %>%
  filter(yr != 1973) %>%
  filter(yr >= 1974) %>%
  filter(yr < 2014) %>% 
  filter(med_sal_ > 0) %>% 
  filter(n > 5) %>%
  drop_na(med_sal_) %>%
  filter(AGYSUB != "TOTL")%>%
  filter(!grepl("NET", agy_full)) %>%
  mutate(logMA_pct = log(ma_pct + 1)) %>%
  mutate(logLST_pct = log(LST_pct + 1)) %>%
  mutate(logAIN_pct = log(AIN_pct + 1)) %>%
  mutate(logPHDpct = log(doc_pct + 1)) %>%
  mutate(logn = log(n)) %>%
  mutate(Environmental = ifelse(agy_typ == "Natural Resources and Environment", 1, 0)) %>%
  mutate(Environmental = ifelse(AGYSUB == "EP00", 2, Environmental)) %>%
  mutate(Environmental = as.factor(Environmental))

sumdf = df %>% 
  group_by(agy_typ) %>%
  summarise("N Agencies" = n(),
            Mean1 = mean(n), SD1 = sd(n),
            Mean2 = mean(med_sal_), SD2 = sd(med_sal_),
            Mean3 = mean(LOSavg), SD3 = sd(LOSavg),
            Mean4 = mean(ma_pct), SD4 = sd(ma_pct),
            Mean5 = mean(AIN_pct), SD5 = sd(AIN_pct),
            Mean6 = mean(LST_pct), SD6 = sd(LST_pct),
            Mean7 = mean(gini, na.rm = T), SD7 = sd(gini, na.rm = T))

sumdf[,2:(ncol(sumdf)-2)] = round(sumdf[,2:(ncol(sumdf)-2)])

sumdf[,c("Mean7", "SD7")] = round(sumdf[,c("Mean7", "SD7")], 2)

colnames(sumdf) = c("Agency Type", "N Agencies", rep(c("Mean", "sd"), 7))

kbl(sumdf, booktabs = T, linesep = "", caption = "Summary Statistics of Capacity Indicators By Agency Type") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  add_header_above(c(" " = 2, "Employees" = 2, "Median Salary" = 2,
                     "Length of\nService" = 2, "% Master's" = 2,
                     "% Employees\nRequisitioned" = 2, "% Employees\nLost" = 2,
                     "Gini" = 2))

```

```{r}

fed = read_csv("fed_agency_capacity_autonomy.csv")

set.seed(9876)
df = fed %>%
  filter(yr != 1973) %>%
  filter(yr >= 1974) %>%
  filter(yr < 2014) %>% 
  filter(med_sal_ > 0) %>% 
  filter(n > 5) %>%
  drop_na(med_sal_) %>%
  filter(AGYSUB != "TOTL")%>%
  filter(!grepl("NET", agy_full)) %>%
  mutate(logMA_pct = log(ma_pct + 1)) %>%
  mutate(logLST_pct = log(LST_pct + 1)) %>%
  mutate(logAIN_pct = log(AIN_pct + 1)) %>%
  mutate(logPHDpct = log(doc_pct + 1)) %>%
  mutate(logn = log(n)) %>%
  mutate(Environmental = ifelse(agy_typ == "Natural Resources and Environment", 1, 0)) %>%
  mutate(Environmental = ifelse(AGYSUB == "EP00", 2, Environmental)) %>%
  mutate(Environmental = as.factor(Environmental))

regressors = c("logn", "logMA_pct", "med_sal_", "logLST_pct", "logAIN_pct", "LOSavg", "gini",
               "n", "ma_pct", "LST_pct", "AIN_pct")

dff = df %>%
  select(regressors) %>%
  mutate_at(regressors, scale) %>%
  drop_na()

dfff = dff %>%
  gather("Variable", "Value", 1:ncol(dff)) %>%
  mutate("logTransform" = as.factor(ifelse(Variable %in% c("logn", "logMA_pct",
                                                  "logLST_pct", "logAIN_pct"), "Yes", "No"))) %>%
  mutate(Variable = ifelse(Variable == "logn", "n", Variable)) %>%
  mutate(Variable = ifelse(Variable == "logAIN_pct", "AIN_pct", Variable)) %>%
  mutate(Variable = ifelse(Variable == "logLST_pct", "LST_pct", Variable)) %>%
  mutate(Variable = ifelse(Variable == "logMA_pct", "ma_pct", Variable)) %>%
  mutate(Variable = ifelse(Variable == "n", "Number of Employees", Variable)) %>%
  mutate(Variable = ifelse(Variable == "AIN_pct", "Percent Employees Gained", Variable)) %>%
  mutate(Variable = ifelse(Variable == "LST_pct", "Percent Employees Lost", Variable)) %>%
  mutate(Variable = ifelse(Variable == "ma_pct", "Percent have Master's Degree", Variable))
  

dfff %>%
  ggplot(aes(x = Value, y = Variable)) +
  stat_density_ridges(aes(fill = logTransform), from = -2, to = 2,
                      alpha = .8, color = "white", panel_scaling = FALSE,
                      rel_min_height = .045) +
  ylab("") +
  xlab("Scaled Value") +
  theme_classic() +
  scale_fill_brewer(type = "qual", palette = 4) +
  ggtitle("Figure 1: Density Plots for Observed Variables")


```

```{r, fig.height=10, fig.width=8}

fed = read_csv("fed_agency_capacity_autonomy.csv")

set.seed(9876)
df = fed %>%
  filter(yr != 1973) %>%
  filter(yr >= 1974) %>%
  filter(yr < 2014) %>% 
  filter(med_sal_ > 0) %>% 
  filter(n > 5) %>%
  drop_na(med_sal_) %>%
  filter(AGYSUB != "TOTL")%>%
  filter(!grepl("NET", agy_full)) %>%
  mutate(logMA_pct = log(ma_pct + 1)) %>%
  mutate(logLST_pct = log(LST_pct + 1)) %>%
  mutate(logAIN_pct = log(AIN_pct + 1)) %>%
  mutate(logPHDpct = log(doc_pct + 1)) %>%
  mutate(logn = log(n)) %>%
  mutate(Environmental = ifelse(agy_typ == "Natural Resources and Environment", 1, 0)) %>%
  mutate(Environmental = ifelse(AGYSUB == "EP00", 2, Environmental)) %>%
  mutate(Environmental = as.factor(Environmental))

df2 = df %>%
  filter(agy_typ == 'Natural Resources and Environment') %>%
  filter(AGYSUB!="EP00")

df3 = df %>%
  filter(AGYSUB == "EP00")


a = ggplot() +
  geom_line(aes(x = yr, y = logn, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = logn, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = logn, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("log(Number of Employees)") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

b = ggplot() +
  geom_line(aes(x = yr, y = med_sal_, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = med_sal_, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = med_sal_, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("Median Salary") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

c = ggplot() +
  geom_line(aes(x = yr, y = LOSavg, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = LOSavg, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = LOSavg, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("Average Length of Service") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

d = ggplot() +
  geom_line(aes(x = yr, y = logMA_pct, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = logMA_pct, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = logMA_pct, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("log(% Have Master's Degree)") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

e = ggplot() +
  geom_line(aes(x = yr, y = logAIN_pct, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = logAIN_pct, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = logAIN_pct, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("log(% Employees Gained From Other Agencies)") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

f = ggplot() +
  geom_line(aes(x = yr, y = logLST_pct, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = logLST_pct, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = logLST_pct, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("log(% Employees Lost to Other Agencies)") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

g = ggplot() +
  geom_line(aes(x = yr, y = gini, group = AGYSUB), data = df, alpha = .4, size = 1, color = "slategray3") +
  geom_line(aes(x = yr, y = gini, group = AGYSUB), data = df2, alpha = .8, size = 1.2, color = "mediumseagreen") +
  geom_line(aes(x = yr, y = gini, group = AGYSUB), data = df3, size = 1.5, color = "orangered1") +
  theme_bw()+ 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_rect(fill = "white")) +
  ggtitle("Gini Coefficient") +
  xlab("") + ylab("") + theme(plot.title = element_text(size=10))

grid.arrange(a, b, c, d, e, f, g, ncol = 2, top = "Figure 2: Observed Trend of Capacity Variables\nGreen = Environmental Agency, Red = EPA ")


```

```{r}

regressors = c("logn", "logMA_pct", "med_sal_", "logLST_pct", "logAIN_pct", "LOSavg", "gini")

cordf = df %>%
  select(regressors) %>%
  drop_na() %>%
  rename(Gini = gini, "Log(Master's %)" = logMA_pct, "Length of Service" = LOSavg,
         "Log(Emp. Gained %)" = logAIN_pct, "Log(Emp. Lost%)" = logLST_pct, "Median Salary" = med_sal_,
         "Log(n Employees)" = logn)

corr = round(cor(cordf), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower",
           outline.col = "white", lab = TRUE, colors = brewer.pal(n = 3, "RdBu")) + ggtitle("Figure 3: Correlation of Observed Capacity Indicators") +
  theme(axis.text.x = element_text(size = 10), axis.text.y = element_text(size = 10))

```

## Empirical Strategy

* Motivation for latent variable (we don't observe capacity, only theoretical indicators)
* Cross Sectional Latent form (should I even include this?):

$$  y_{i,t,j} \sim \mathcal{N}(\mu_{i,t,j}, \sigma_{t, j} ^2) $$
$$  \mu_{i, t, j} = \alpha_{t,j} + \beta_{t,j} * Z_{i,t} $$

Identification restriction:

$$  Z_{i,t} \sim \mathcal{N}(0, 1) $$

* This specification does not account for covariance between time points

* Jackman: Dynamic measurement model in Bayesian framework allows inference over a longer longitudinal dimension.

$$  \mathbf{y}_{i,t} = \mathbf{F}_{i, t} \mathbf{Z_{i,t}} + \mathbf{W}_{i,t}\boldsymbol\delta_{i,t} + \boldsymbol{\epsilon}_{i,t} $$

* Latent variable evolves:

$$  \mathbf{Z}_{i,t} = \mathbf{G}_{i,t}\mathbf{Z_{i, t-1}} + \mathbf{A}_t \mathbf{x}_{i,t} + \mathbf{u}_{i,t} $$

* Random Walk model sets $\mathbf{G}_{i,t} = \mathbf{I}, \forall i,t$
* Say used Kalman filter, marginal likelihood with respect to the latent variable. 

## Results

# References